---
title: "2023-01-27-torR-exp4-bioscreen"
author: "Amy Schmid"
date: "1/27/2023"
output: html_document
---


```{r}
# fresh start
rm(list = ls()) 
```

```{r}
# load required libraries
library(plyr)
library(tidyr) 
library(dplyr)
library(tibble)
library(viridis)
library(ggplot2)
library(tinytex)
library(latexpdf)
library(data.table)
library(growthrates)
library(gridExtra)
library(growthcurver)
library(factoextra)
library(Rfast)
library(rstatix)
library(ggpubr)
library(plotrix)
```

```{r}
##################################################################################################### SETTINGS
# project folder
folder <- '~/growthdata/'
# folder with meta
foldermeta <- '~/growthdata/meta/'
# metafile name
meta <- '221108dtxRtitration_metadata-AKS.csv'
# folder with data
folderdata <- '~/growthdata/data/'
# datafile name
data <- '221108dtxRtitration_data-AKS.csv'
# wells to be removed based on observation (do not remove = 1000)
remove1<- 159
remove2<- 1000
remove3<- 1000
remove4<- 1000
remove5<- 1000
remove6<- 1000
remove7<- 1000
remove8<- 1000
remove9<- 1000
remove10<- 1000
remove11<- 1000
remove12<- 1000
# blank wells (not a blank = 1000)
blank1<- 1
blank2<- 2
blank3<- 3
blank4<- 4


```

```{r}

##################################################################################################### RAW DATA
# load data
## read in metafile
mt <- read.csv(paste(foldermeta, meta, sep = ''),
               sep = ',',
               header = TRUE)
mt0 <- mt[-1, -(1:2)]
## read in datafile
dt <- read.csv(paste(folderdata, data, sep = ''),
               sep = ',',
               header = FALSE)
dt0 <- dt[-(1:2), -2]
colnames(dt0) <- c('time', dt0[1, -1])
dt0 <- dt0[-1, ]

# remove rows based on observation in meta
mt0 <- mt0[!(mt0$variable == remove1 |
               mt0$variable == remove2 |
               mt0$variable == remove3 |
               mt0$variable == remove4 |
               mt0$variable == remove5 |
               mt0$variable == remove6 |
               mt0$variable == remove7 |
               mt0$variable == remove8 |
               mt0$variable == remove9 |
               mt0$variable == remove10 |
               mt0$variable == remove11 |
               mt0$variable == remove12),]

# remove rows based on observation in data
dt0 <- dt0 %>% select(-contains(as.character(remove1)))
dt0 <- dt0 %>% select(-contains(as.character(remove2)))
dt0 <- dt0 %>% select(-contains(as.character(remove3)))
```

```{r}
# extract hours, mins, secs
hou <-
  as.numeric(lapply(strsplit(as.character(dt0$time), ':'), '[[', 1))
min <-
  as.numeric(lapply(strsplit(as.character(dt0$time), ':'), '[[', 2))
sec <-
  as.numeric(lapply(strsplit(as.character(dt0$time), ':'), '[[', 3))
## set total time to hours
dt0$time <- (hou) + min / 60 + (sec / 3600)

# melt the dataframe to one list
m_dt <- melt(dt0, id = 'time')
## merge
mtdt <- merge(m_dt, mt0, by = 'variable')
mtdt$environment <- ifelse(mtdt$condition=='',as.character(mtdt$media),paste(mtdt$media,'+',mtdt$condition,sep=''))
mtdt$ID <- ifelse(mtdt$condition=='',as.character(mtdt$name),paste(mtdt$name,' (',mtdt$condition,')',sep=''))
well <- paste(mtdt$variable, mtdt$strain, sep = ' ')
mtdt <- cbind(well, mtdt)
## remove rows without ID
c_mtdt <- mtdt[!(mtdt$ID == ''), ]

# time cutoff
t_mtdt <- c_mtdt[!(c_mtdt$time <= 0.5 | c_mtdt$time >= 65), ]
t_mtdt[500:510,]
## fix delta symbol
#t_mtdt$ID<- gsub('_','\u0394',t_mtdt$ID)
#t_mtdt$name<- gsub('_','\u0394',t_mtdt$name)

## select blanks (three tech reps each) 
#blanks<- t_mtdt[(t_mtdt$variable == 1 |
        #           t_mtdt$variable == 2 |
         #          t_mtdt$variable == 3 |
          #         t_mtdt$variable == 4 |
          #         t_mtdt$variable ==  5 |
          #         t_mtdt$variable == 6 |
          #         t_mtdt$variable == 7 |
          #         t_mtdt$variable == 8 |
          #         t_mtdt$variable == 9 |
          #         t_mtdt$variable == 10 |
          #         t_mtdt$variable == 11 |
          #         t_mtdt$variable == 12), ]
#blanks <- filter(t_mtdt, strain == "blank")

## summarize all values (time, strain, media, condition, biorep)
tm_mtdt<- ddply(t_mtdt, c('time', 'ID', 'strain', 'name', 'media', 'condition', 'environment', 'biorep'),
               summarize,
              value_mn = mean(value),
               value_sd = sd(value))

```

```{r}
# set baseline (blank wells)
## take minimal value of each blank per condition
#blank_all<- ddply(blanks, c('environment','time'),
 #             summarize,
  #            blank=min(value)-0.01)

## add correct blank values to each row
#bu_mtdt<- merge(t_mtdt,blank_all,c('environment','time'))
## reorganize columns
#bu_mtdt$value_mn<- bu_mtdt$value - bu_mtdt$blank  
#bu_mtdt$blank<- NULL
## summarize all blank corrected values (time, strain, media, condition, biorep)
#bm_mtdt<- ddply(bu_mtdt, c('time', 'ID', 'strain', 'name', 'media', 'condition', 'environment', 'biorep'),
#                summarize,
#                value_mn = mean(value),
#                value_sd = sd(value))
## remove blank rows
#b_mtdt <- bu_mtdt[!bu_mtdt$name=='blank',]
#bm_mtdt <- bm_mtdt[!bm_mtdt$name=='blank',]
#min(bm_mtdt$value_mn)
#min(bm_mtdt$value_mn)

```

```{r}
##################################################################################################### PLOT BLANKS
# all growth curves
# folder for plots
dir.create(paste(folder,'plots',sep=''))
setwd(paste(folder,'plots',sep=''))
# blanks
p_blanks<- 
  ggplot(blanks, aes(x = time, y = value, color = environment)) +
  ggtitle('Blank curves') +
  xlab('time (h)') + 
  ylab(bquote(bold(paste('Log'['2'],'(Absorbance at 600nm)')))) +
  geom_line() +
  scale_y_continuous(trans = 'log2') +
  scale_color_viridis('condition', discrete = TRUE) +
  theme(#legend.position = 'none',
    strip.text = element_text(face = "italic",size = 8),
    plot.title = element_text(face='bold', color='#000000', size=12, angle=0),
    axis.title.x = element_text(face='bold', color='#000000', size=10, angle=0),
    axis.title.y = element_text(face='bold', color='#000000', size=10, angle=90),
    legend.title = element_text(face='bold', color='#000000', size=10, angle=0),
    axis.text.y = element_text(face='plain', color='#000000', size=8, angle=0),
    axis.text.x = element_text(face='plain', color='#000000', size=8, angle=0),
    legend.text = element_text(face='plain', color='#000000', size=10, angle=0))

ggsave(p_blanks,
       file = paste(
         format(Sys.time(), '%Y%m%d'),
         '_',
         'plot_blanks_exp4',
         #'.jpeg',
         '.pdf',
         sep = ''),
       width = 10, height = 8, units = "in", device='pdf')
```

```{r}
#blanks show increase in OD for some conditions. remove blanks & empty wells and do not use blank-corrected data.
tm_mtdt <- tm_mtdt[!(tm_mtdt$name == "blank"), ]
#t_mtdt <- t_mtdt[!(t_mtdt$name == "empty"), ]
t_mtdt <- t_mtdt[!(t_mtdt$name == "blank"), ]
#t_mtdt <- t_mtdt[!(t_mtdt$name == "empty"), ]
```

```{r}
##################################################################################################### PLOT RAW
# folder for plots
dir.create(paste(folder,'plots',sep=''))
setwd(paste(folder,'plots',sep=''))

# growthcurves - one plot - well
p_well <- 
  ggplot(t_mtdt, aes(x = time, y = value, color = well)) +
  ggtitle('Growthcurves each well') +
  xlab('time (h)') + 
 ylab(bquote(bold(paste('Log'['2'],'(Absorbance at 600nm)')))) +
  geom_line() +
  scale_y_continuous(trans = 'log2') +
  scale_color_viridis('well', discrete = TRUE) +
  theme(legend.position = 'none',
        strip.text = element_text(face = "italic",size = 8),
        plot.title = element_text(face='bold', color='#000000', size=12, angle=0),
        axis.title.x = element_text(face='bold', color='#000000', size=10, angle=0),
        axis.title.y = element_text(face='bold', color='#000000', size=10, angle=90),
        legend.title = element_text(face='bold', color='#000000', size=10, angle=0),
        axis.text.y = element_text(face='plain', color='#000000', size=8, angle=0),
        axis.text.x = element_text(face='plain', color='#000000', size=8, angle=0),
        legend.text = element_text(face='plain', color='#000000', size=10, angle=0))

ggsave(p_well, file = paste(
  format(Sys.time(), '%Y%m%d'),
  '_',
  'plot_well_all_exp4',
 # '.jpeg',
 '.pdf',
  sep = ''),
  width = 10, height = 8, units = "in", device='pdf')
```

```{r}
# growthcurves - one plot - ID
p_id <- 
  ggplot(t_mtdt, aes(x = time, y = value, color = ID)) +
  ggtitle('Growthcurves each strain') +
  xlab('time (h)') + 
  ylab(bquote(bold(paste('Log'['2'],'(Absorbance at 600nm)')))) +
  geom_point() +
  scale_y_continuous(trans = 'log2') +
  scale_color_manual('strain description', values = rainbow(25)) +
  theme(#legend.position = 'none',
    strip.text = element_text(face = "italic",size = 8),
    plot.title = element_text(face='bold', color='#000000', size=12, angle=0),
    axis.title.x = element_text(face='bold', color='#000000', size=10, angle=0),
    axis.title.y = element_text(face='bold', color='#000000', size=10, angle=90),
    legend.title = element_text(face='bold', color='#000000', size=10, angle=0),
    #legend.position = "none",
    axis.text.y = element_text(face='plain', color='#000000', size=8, angle=0),
    axis.text.x = element_text(face='plain', color='#000000', size=8, angle=0),
    legend.text = element_text(face='italic', color='#000000', size=10, angle=0))

ggsave(p_id, file = paste(
  format(Sys.time(), '%Y%m%d'),
  '_',
  'plot_ID_all_exp4',
  #'.jpeg',
  '.pdf',
  sep = ''),
  width = 10, height = 8, units = "in", device='pdf')

#experiments with YPC medium need to be plotted separately from those with PR - different time scale and growth rate.
```

```{r}
# growthcurves - grid
p_gri <-
  ggplot(t_mtdt, aes(x = time, y = value)) +
  ggtitle('Growthcurves each well') +
  xlab('time (h)') + 
 ylab(bquote(bold(paste('Log'['2'],'(Absorbance at 600nm)')))) +
  geom_line() +
  scale_y_continuous(trans = 'log2') +
  facet_wrap(~ well) +
  theme(#legend.position = 'none',
    strip.text = element_text(face = "plain",size = 8),
    plot.title = element_text(face='bold', color='#000000', size=12, angle=0),
    axis.title.x = element_text(face='bold', color='#000000', size=8, angle=0),
    axis.title.y = element_text(face='bold', color='#000000', size=8, angle=90),
    legend.title = element_text(face='bold', color='#000000', size=8, angle=0),
    axis.text.y = element_text(face='plain', color='#000000', size=6, angle=0),
    axis.text.x = element_text(face='plain', color='#000000', size=6, angle=0),
    legend.text = element_text(face='plain', color='#000000', size=8, angle=0))

ggsave(p_gri, file = paste(
  format(Sys.time(), '%Y%m%d'),
  '_',
  'plot_grid_all_exp4',
  '.jpeg',
  sep = ''),
  width = 15, height = 10, units = "in", device='jpeg')



#not a useful plot layout, can't see the plots for the labels. need multipage pdf.
```
# Plot all curves average and 95% conf interval, first grouped by strain
```{r}

##################################################################################################### PLOT STATS
# folder for plots
dir.create(paste(folder,'plots',sep=''))
setwd(paste(folder,'plots',sep=''))
```

```{r}
# 95% confidence interval function
conf_int95 <- function(data) {
  n <- length(data)
  error <- qt(0.975, df = n - 1) * sd(data) / sqrt(n)
  return(error)
}
```

```{r}
stats <- tm_mtdt %>%
  group_by(ID, strain, name, environment, time) %>%
  dplyr::summarise(
    reps = length(value_mn),
    average = mean(value_mn),
    CI95 = conf_int95(value_mn)) #%>%
  #filter(!strain == '')



# growth curves - average strain
suppressWarnings({
p_ave_str <-
ggplot(stats, aes(x = time, y = average, color = environment)) +
  ggtitle('Growthcurves averaged each condition and separated by strain') +
  xlab('time (h)') + 
 ylab(bquote(bold(paste('Log'['2'],'(Absorbance at 600nm)')))) +
  geom_ribbon(
    aes(
      ymin = average - CI95,
      ymax = average + CI95,
      fill = environment), color = NA, alpha = 0.3) +
  geom_line() +
  scale_y_continuous(trans = 'log2') +
  facet_grid(name ~ .) +
  scale_fill_manual('condition', values = rainbow(8)) +
  scale_color_manual('condition', values = rainbow(8)) +
  theme(#legend.position = 'none',
    strip.text = element_text(face = "italic",size = 7),
    plot.title = element_text(face='bold', color='#000000', size=12, angle=0),
    axis.title.x = element_text(face='bold', color='#000000', size=10, angle=0),
    axis.title.y = element_text(face='bold', color='#000000', size=10, angle=90),
    legend.title = element_text(face='bold', color='#000000', size=10, angle=0),
    axis.text.y = element_text(face='plain', color='#000000', size=8, angle=0),
    axis.text.x = element_text(face='plain', color='#000000', size=8, angle=0),
    legend.text = element_text(face='plain', color='#000000', size=10, angle=0))
   #legend.position = "none")

ggsave(p_ave_str,
       file = paste(
         format(Sys.time(), '%Y%m%d'),
         '_',
         'plot_average_strain_exp4',
         '.pdf',
         sep = ''),
       width = 4, height =6, units = "in", device='pdf')

})

```

#Plot all average curves +/- 95% conf interval, now grouped by condition
```{r}
# growth curves - average condition
p_ave_con <-
ggplot(stats, aes(x = time, y = average, color = name)) +
  ggtitle('Growthcurves averaged each strain and separated by condition') +
  xlab('time (h)') + 
 ylab(bquote(bold(paste('Log'['2'],'(Absorbance at 600nm)')))) +
  geom_ribbon(
    aes(
      ymin = average - CI95,
      ymax = average + CI95,
      fill = name), color = NA, alpha = 0.3) +
  geom_line() +
  scale_y_continuous(trans = 'log2') +
  facet_grid(environment ~ .) +
  scale_fill_manual('Strain name', values = rainbow (5)) +
  scale_color_manual('Strain name', values = rainbow(5)) +
  theme(#legend.position = 'none',
    strip.text = element_text(face = "italic",size = 10),
    plot.title = element_text(face='bold', color='#000000', size=12, angle=0),
    axis.title.x = element_text(face='bold', color='#000000', size=10, angle=0),
    axis.title.y = element_text(face='bold', color='#000000', size=10, angle=90),
    legend.title = element_text(face='bold', color='#000000', size=10, angle=0),
    axis.text.y = element_text(face='plain', color='#000000', size=8, angle=0),
    axis.text.x = element_text(face='plain', color='#000000', size=8, angle=0),
    legend.text = element_text(face='italic', color='#000000', size=10, angle=0))

ggsave(p_ave_con,
      file = paste(
        format(Sys.time(), '%Y%m%d'),
        '_',
        'plot_average_condition_exp4',
        '.pdf',
        sep = ''),
      width = 4, height = 6, units = "in", device='pdf')
```





```{r}

##################################################################################################### MUMAX Growthrates
# folder for plots
dir.create(paste(folder,'plots',sep=''))
setwd(paste(folder,'plots',sep=''))

# extract mumax for plotting fit
pretty_spline_fits <-
  all_splines(value ~ time | ID + biorep,
              data = t_mtdt,
              spar = 0.5)
# extract mumax for table
all_spline_fits <-
  all_splines(value ~ time | ID + biorep + strain + name + environment,
              data = t_mtdt,
              spar = 0.5)

#plot spline fit
jpeg(paste(
  format(Sys.time(), '%Y%m%d'),
  '_',
  'plot_splines_exp4',
  '.jpeg',sep=''),
  width = 1188, 
  height = 880)
par(mfrow = c(8, 7))
par(mar = c(2, 2, 1.5, 1.25))
plot(pretty_spline_fits, log='y')
dev.off()

#spline fits are not good - troR especially the spline fit only captures the first couple hours of the curve.

# store results in dataframe
mumax<- results(all_spline_fits)
row.names(mumax)<- NULL

# remove blanks
mumax <- mumax[!mumax$strain=='blank',]
mumax.noHA <- mumax[(!mumax$strain == 'blank' & !mumax$name == 'trmB-HA' & !mumax$name == 'parent+ptrmB' & !mumax$name == 'parent+ptrmB-HA' & !mumax$name == 'ΔtrmB+ptrmB-HA'), ]

# remove mumax with rsquared below 0.95
mumax$mumax <- ifelse(mumax$r2 < 0.95, 0, mumax$mumax)
#mumax.noHA <-  ifelse(mumax.noHA$r2 < 0.95, 0, mumax.noHA$mumax)
# calculate mean mumax
mumean<- ddply(mumax,c('ID','strain','name','environment'), 
               summarize,
               OD_mn=mean(y0),
               OD_sd=sd(y0),
               mumax_mn=mean(mumax),
               mumax_sd=sd(mumax))

```



#assign colors based on strains, reorder
```{r}
##trmB only = 'yellow'
##trmB-HA&trmB = 'mediumseagreen'
##trmB-HA only = 'darkslategray4'
##neither = 'darkorchid4'
#modcol<- c('trmB only','trmB only','trmB only','trmB only','trmB only',
 #          'trmB only','trmB-HA&trmB','trmB-HA&trmB','trmB-HA only','trmB-HA only',
  #         'neither','neither','neither','neither','trmB only',
  #         'trmB only','trmB-HA only','trmB-HA only')



#mumean<- cbind(modcol,mumean)
#mumean.noHA <- cbind (modcol.noHA, mumean.noHA)

#reorder
#mumean<- mumean[order(mumean$mumax_mn),]
#mumean$ID <- factor(mumean$ID, levels = mumean$ID[order(mumean$mumax_mn)])

#mumean.noHA<- mumean.noHA[order(mumean.noHA$mumax_mn),]
#mumean.noHA$ID <- factor(mumean.noHA$ID, levels = mumean.noHA$ID[order(mumean.noHA$mumax_mn)])
```

```{r}
# growthcurves - bars
p_mu <- 
  ggplot(data = mumean, aes(x = ID, y = mumax_mn, fill = strain)) +
  ggtitle('Mean mumax of bioreps comparison') +
  xlab('Strain name (condition)') + 
  ylab('[mumax]')+
  geom_bar(stat='identity') +
  geom_errorbar(aes(ymin=mumax_mn-mumax_sd, ymax=mumax_mn+mumax_sd), width=.2,
              position=position_dodge(.9))+
  coord_flip() +
  scale_fill_manual(values = rainbow(8)) +
  theme(#legend.position = 'none',
    plot.title = element_text(face='bold', color='#000000', size=12, angle=0),
    axis.title.x = element_text(face='bold', color='#000000', size=10, angle=0),
    axis.title.y = element_text(face='bold', color='#000000', size=10, angle=90),
    legend.title = element_text(face='bold', color='#000000', size=10, angle=0),
    axis.text.y = element_text(face='italic', color='#000000', size=8, angle=0),
    axis.text.x = element_text(face='plain', color='#000000', size=8, angle=0),
    legend.text = element_text(face='italic', color='#000000', size=10, angle=0))

ggsave(p_mu,
       file = paste(
         format(Sys.time(), '%Y%m%d'),
         '_',
         'plot_mumax_exp4',
         '.jpeg',
        
         sep = ''),
       width = 10, height = 8, units = "in", device='jpeg')
```






```{r}
# growthcurves - bars
p_mu <- 
  ggplot(data = mumean, aes(x = ID, y = mumax_mn, fill = modcol)) +
  ggtitle('Mean mumax of bioreps comparison') +
  xlab('Strain name (condition)') + 
  ylab('[mumax]')+
  geom_bar(stat='identity') +
  geom_errorbar(aes(ymin=mumax_mn-mumax_sd, ymax=mumax_mn+mumax_sd), width=.2,
                position=position_dodge(.9))+
  coord_flip() +
  scale_fill_manual('available trmB genes',values=c('blank'='white','trmB only'='yellow3','trmB-HA&trmB'='mediumseagreen',
                                                    'neither' = 'darkorchid4','trmB-HA only' = 'darkslategray4')) +
  theme(#legend.position = 'none',
    plot.title = element_text(face='bold', color='#000000', size=12, angle=0),
    axis.title.x = element_text(face='bold', color='#000000', size=10, angle=0),
    axis.title.y = element_text(face='bold', color='#000000', size=10, angle=90),
    legend.title = element_text(face='bold', color='#000000', size=10, angle=0),
    axis.text.y = element_text(face='italic', color='#000000', size=8, angle=0),
    axis.text.x = element_text(face='plain', color='#000000', size=8, angle=0),
    legend.text = element_text(face='italic', color='#000000', size=10, angle=0))

ggsave(p_mu,
       file = paste(
         format(Sys.time(), '%Y%m%d'),
         '_',
         'plot_mumax_minb',
         '.jpeg',
         sep = ''),
       width = 10, height = 8, units = "in", device='jpeg')
```

## calculate growth metrics using growthcurver package
```{r}
# Let's create an output data frame to store the results in. We'll create it so that it is the right size (it's faster this way!), but leave it empty.
num_analyses <- length(names(dt0)) - 1
d_gc <- data.frame(
  sample = character(num_analyses),
  k = numeric(num_analyses),
  n0 = numeric(num_analyses),
  r = numeric(num_analyses),
  t_mid = numeric(num_analyses),
  t_gen = numeric(num_analyses),
  auc_l = numeric(num_analyses),
  auc_e = numeric(num_analyses),
  sigma = numeric(num_analyses),
  stringsAsFactors = FALSE
)
```

```{r}
# Truncate or trim the input data to observations occuring in the first 60 hours to fit accurately the exponential growth phase.
# Remember that the times in these sample data are reported in hours. To use
# minutes (or to trim at a different time), change the next line of code.
# For example, if you still would like to trim at 50 hours, but your time data
# are reported in minutes use: trim_at_time <- 20 * 60
trim_at_time <- 65
#name the experiment
expt <- 'exp4_221108dtxRtitration'

# Now, loop through all of the columns in the data frame. For each column,
# run Growthcurver, save the most useful metrics in the output data frame,
# and make a plot of all the growth curve data and their best fits.

# First, create a plot for each of the wells in the 96-well plate.
# Uncomment the next line to save the plots from your 96-well plate to a
# pdf file in the working directory.
pdf(paste(expt, "_growthcurver_fits.pdf", sep = ""), height = 8.5, width = 11)
par(mfcol = c(8, 12))
par(mar = c(0.25, 0.25, 0.25, 0.25))
y_lim_max <- max(dt0[, setdiff(names(dt0), "time")], na.rm = T) - min(dt0[, setdiff(names(dt0), "time")], na.rm = T)

n <- 1 # keeps track of the current row in the output data frame
for (col_name in names(dt0)) {

  # Don't process the column called "time".
  # It contains time and not absorbance data.
  if (col_name != "time") {

    # Create a temporary data frame that contains just the time and current col
    d_loop <- dt0[c("time", col_name)]
    # Do the background correction.
    # Background correction option 1: subtract the minimum value in a column
    #                                 from all measurements in that column
    # min_value <- min(d_loop[, col_name])
    # d_loop[, col_name] <- d_loop[, col_name] - min_value
    # Background correction option 2: subtract the mean value of blank wells
    #                                 over the course the experiment
    #                                 (Replace B2, D8, G11 with the column
    #                                  names of your media-only wells)
    # d_loop[col_name] <- d_loop[col_name] - d$blank
    # Now, call Growthcurver to calculate the metrics using SummarizeGrowth
    gc_fit <- SummarizeGrowth(
      data_t = d_loop$time,
      data_n = d_loop[col_name],
      t_trim = trim_at_time,
      bg_correct = "none"
    )
    # print(gc_fit$vals[[16]])

    if (gc_fit$vals[[16]] == "cannot fit data") {
      d_gc$sample[n] <- col_name
      n <- n + 1
    } else {
      # Now, add the metrics from this column to the next row (n) in the
      # output data frame, and increment the row counter (n)
      d_gc$sample[n] <- col_name
      d_gc[n, 2:9] <- c(
        gc_fit$vals$k,
        gc_fit$vals$n0,
        gc_fit$vals$r,
        gc_fit$vals$t_mid,
        gc_fit$vals$t_gen,
        gc_fit$vals$auc_l,
        gc_fit$vals$auc_e,
        gc_fit$vals$sigma
      )

      n <- n + 1
      # Finally, plot the raw data and the fitted curve
      # Here, I'll just print some of the data points to keep the file size smaller
      n_obs <- length(gc_fit$data$t)
      idx_to_plot <- 1:20 / 20 * n_obs
      plot(gc_fit$data$t[idx_to_plot], gc_fit$data$N[idx_to_plot],
        pch = 20,
        xlim = c(0, trim_at_time),
        ylim = c(0, y_lim_max),
        cex = 0.6, xaxt = "n", yaxt = "n"
      )
      text(x = trim_at_time / 4, y = y_lim_max, labels = col_name, pos = 1)
      lines(gc_fit$data$t, predict(gc_fit$model), col = "red")
    }
  }
}
# Uncomment the next line to save the plots from your 96-well plate to a file
dev.off()
d_gc
```

#graph fitted curves:
## First get fitted curves and params
```{r, message=F}
fitted_curves <- list()

# Now, loop through all of the columns in the data frame. For each column, run Growthcurver, save the most useful metrics in the output data frame.
n <- 1 # keeps track of the current row in the output data frame
for (col_name in names(dt0)) {

  # Don't process the column called "time".
  # It contains time and not absorbance data.
  if (col_name != "time") {

    # Create a temporary data frame that contains just the time and current col
    d_loop <- dt0[c("time", col_name)]
    # Now, call Growthcurver to calculate the metrics using SummarizeGrowth
    gc_fit <- SummarizeGrowth(
      data_t = d_loop$time,
      data_n = d_loop[col_name],
      t_trim = trim_at_time,
      bg_correct = "none"
    )

    if (gc_fit$vals[[16]] == "cannot fit data") {
      d_gc$sample[n] <- col_name
      n <- n + 1
    } else {
      # Now, add the metrics from this column to the next row (n) in the
      # output data frame, and increment the row counter (n)
      fitted_curves[[n]] <- predict(gc_fit$model)
      names(fitted_curves)[n] <- col_name
      n <- n + 1
    }
  }
}

tmp <- tibble(got = fitted_curves)
tmp <- Filter(length, fitted_curves)
fits <- as.data.frame(tmp, col.names = names(tmp))
colnames(fits) <- names(tmp)
fits <- cbind("time" = dt0$time[dt0$time < trim_at_time], fits)

# re-zero lowest value
mins <- Rfast::colMins(as.matrix(fits), value = T) # get column minimums
fits <- sweep(fits, 2, mins, "-") # subtract smallest value for each column

fits
```
## Now plot and export fitted curve graphs with CI95 as shaded error region
```{r, fig.width=6, fig.height=4}
# Convert data from wide to long format
m_dt <- reshape2::melt(fits, id = "time")
mtdt <- merge(m_dt, mt0, by = "variable")
mtdt$environment <- ifelse(mtdt$condition=='',as.character(mtdt$media),paste(mtdt$media,'+',mtdt$condition,sep=''))
mtdt$ID <- ifelse(mtdt$condition=='',as.character(mtdt$name),paste(mtdt$name,mtdt$condition,sep=''))
#well <- paste(mtdt$variable, mtdt$strain, sep = ' ')
well <- paste(mtdt$variable, mtdt$ID, sep = " ")
mtdt <- cbind(well, mtdt)

t_mtdt2 <- mtdt[!(mtdt$ID == ""), ]
t_mtdt2 <- mtdt[!(mtdt$name == "blank"), ]


# add 1 to all values for log transformation
stats2 <- t_mtdt2 %>%
  group_by(ID, strain, name, condition, time) %>%
  dplyr::summarise(
    reps = length(value),
    average = mean(log10(value + 1)),
    sd = sd(log10(value + 1)),
    sem = std.error(log10(value+1)),
    CI95 = conf_int95(log10(value + 1))
  )

#export data for compilation across experiments
write.csv (stats2, file = "exp4_fitcurves.csv")
```

```{r}
stats2$facet = factor(stats2$condition, levels = c("(--Fe)", "(-Fe)", "(1Fe)", "(10Fe)", "(100Fe)", "(250Fe)"))

stats2 %>%
  ggplot(., aes(x = time, y = average, color = name)) +
  xlab("time (h)") +
  ylab("log10(Absorbance at 600nm)") +
  geom_line(size = 1) +
  geom_ribbon(aes(ymin = average - sem, ymax = average + sem, fill = name, color = NULL), alpha = 0.2) +
  facet_wrap(~facet) +
  #facet_grid(name~condition) +
  scale_color_manual('strain description', values = rainbow(5)) +
  theme_classic() + 
  
  theme(
    plot.title = element_text(face = "bold", color = "#000000", size = 12, angle = 0),
    axis.title.x = element_text(face = "bold", color = "#000000", size = 10, angle = 0),
    axis.title.y = element_text(face = "bold", color = "#000000", size = 10, angle = 90),
    legend.title = element_blank(), legend.position = c(0.85, 0.15),
    axis.text.y = element_text(face = "plain", color = "#000000", size = 10, angle = 0),
    axis.text.x = element_text(face = "plain", color = "#000000", size = 10, angle = 0),
    legend.text = element_text(face = "plain", color = "#000000", size = 10, angle = 0)
  ) -> fitted.data.sem
plot(fitted.data.sem)

pdf(paste(expt, "_fitcurves_CI95_plot.pdf", sep = ""))
plot(fitted.data.sem)
dev.off()
```

##  plot and export fitted curve graphs with sd as shaded error region
```{r}
stats2 %>%
  ggplot(., aes(x = time, y = average, color = name)) +
  xlab("time (h)") +
  ylab("log10(Absorbance at 600nm)") +
  geom_line(size = 1) +
  geom_ribbon(aes(ymin = average - sd, ymax = average + sd, fill = name, color = NULL), alpha = 0.2) +
  facet_wrap(~facet, nrow = 1) +
  
  theme_classic() +
   scale_color_manual('strain description', values = rainbow(5)) +
  theme(
    plot.title = element_text(face = "bold", color = "#000000", size = 12, angle = 0),
    axis.title.x = element_text(face = "bold", color = "#000000", size = 10, angle = 0),
    axis.title.y = element_text(face = "bold", color = "#000000", size = 10, angle = 90),
    legend.title = element_blank(), legend.position = c(0.85, 0.15),
    axis.text.y = element_text(face = "plain", color = "#000000", size = 10, angle = 0),
    axis.text.x = element_text(face = "plain", color = "#000000", size = 10, angle = 0),
    legend.text = element_text(face = "plain", color = "#000000", size = 10, angle = 0)
  ) -> fitted.data.sd
plot(fitted.data.sd)

pdf(paste(expt, "_fitcurves_sd_plot.pdf", sep = ""), height = 4, width = 16)
plot(fitted.data.sd)
dev.off()

#shaded region for arsR overwhelms all because insufficient data to calculate CI95. If CI95 will be the metric, will need to exclude arsR


```
## plot without arsR, CI95
```{r fig.width=6, fig.height=4}

t_mtdt3 <- t_mtdt2[!(t_mtdt2$strain == "AKS338"), ]


# add 1 to all values for log transformation
stats4 <- t_mtdt3 %>%
  group_by(ID, strain, name, condition, time) %>%
  dplyr::summarise(
    reps = length(value),
    average = mean(log10(value + 1)),
    CI95 = conf_int95(log10(value + 1)),
             sd = sd(log10(value + 1)         )
  )

stats4$facet = factor(stats4$condition, levels = c("(--Fe)", "(-Fe)", "(1Fe)", "(10Fe)", "(100Fe)", "(250Fe)"))

stats4 %>%
  ggplot(., aes(x = time, y = average, color = name)) +
  xlab("time (h)") +
  ylab("log10(Absorbance at 600nm)") +
  geom_line(size = 1) +
  geom_ribbon(aes(ymin = average - CI95, ymax = average + CI95, fill = name, color = NULL), alpha = 0.2) +
  facet_wrap(~facet) +
  theme_classic() +
  
  theme(
    plot.title = element_text(face = "bold", color = "#000000", size = 12, angle = 0),
    axis.title.x = element_text(face = "bold", color = "#000000", size = 10, angle = 0),
    axis.title.y = element_text(face = "bold", color = "#000000", size = 10, angle = 90),
    legend.title = element_blank(), legend.position = c(0.85, 0.15),
    axis.text.y = element_text(face = "plain", color = "#000000", size = 10, angle = 0),
    axis.text.x = element_text(face = "plain", color = "#000000", size = 10, angle = 0),
    legend.text = element_text(face = "plain", color = "#000000", size = 10, angle = 0)
  ) -> fitted.data
plot(fitted.data)

#pdf(paste(expt, "_fitcurves_plot.pdf", sep = ""))
plot(fitted.data)
#dev.off()
#fitted curves look satisfactory for continuing with analysis of AUC diffs
```
##  plot and export fitted curve graphs with sd as shaded error region, pyrE titration only
```{r}
stats2.titr <- filter (stats2, name == "pyrE")
cond.cols <- c(`(--Fe)` = "bisque",  `(-Fe)` = "orangered2", `(10Fe)` = "orangered3", `(100Fe)` = "orangered4", `(250Fe)` = "black")
stats2.titr$facet = factor(stats2.titr$condition, levels = c("(--Fe)", "(-Fe)", "(10Fe)", "(100Fe)", "(250Fe)"))

stats2.titr %>%
  ggplot(., aes(x = time, y = average, color = condition)) +
  xlab("time (h)") +
  ylab("log10(Absorbance at 600nm)") +
  geom_line(size = 1) +
 # geom_ribbon(aes(ymin = average - sem, ymax = average + sem, fill = name, color = NULL), alpha = 0.2) +
 # facet_wrap(~facet, nrow = 1) +
  
  theme_classic() +
  # scale_color_manual('strain description', values = rainbow(5)) +
  scale_color_manual ('condition', values = cond.cols) +
  scale_fill_manual("condition", values = cond.cols) +
  theme(
    plot.title = element_text(face = "bold", color = "#000000", size = 12, angle = 0),
    axis.title.x = element_text(face = "bold", color = "#000000", size = 10, angle = 0),
    axis.title.y = element_text(face = "bold", color = "#000000", size = 10, angle = 90),
    legend.title = element_blank(), legend.position = c(0.85, 0.15),
    axis.text.y = element_text(face = "plain", color = "#000000", size = 10, angle = 0),
    axis.text.x = element_text(face = "plain", color = "#000000", size = 10, angle = 0),
    legend.text = element_text(face = "plain", color = "#000000", size = 10, angle = 0)
  ) -> fitted.data.sem
plot(fitted.data.sem)

#pdf(paste(expt, "_fitcurves_sem_pyre-titr-plot.pdf", sep = ""), height = 4, width = 6)
plot(fitted.data.sem)
#dev.off()

write.csv (stats2.titr, file = "exp4_fitcurves_parent.csv")
```

## plot without arsR, sd
```{r}



stats4 %>%
  ggplot(., aes(x = time, y = average, color = name)) +
  xlab("time (h)") +
  ylab("log10(Absorbance at 600nm)") +
  geom_line(size = 1) +
  geom_ribbon(aes(ymin = average - sd, ymax = average + sd, fill = name, color = NULL), alpha = 0.2) +
  facet_wrap(~facet, nrow=1) +
  theme_classic() +
  
  theme(
    plot.title = element_text(face = "bold", color = "#000000", size = 12, angle = 0),
    axis.title.x = element_text(face = "bold", color = "#000000", size = 10, angle = 0),
    axis.title.y = element_text(face = "bold", color = "#000000", size = 10, angle = 90),
    legend.title = element_blank(), legend.position = c(0.85, 0.15),
    axis.text.y = element_text(face = "plain", color = "#000000", size = 10, angle = 0),
    axis.text.x = element_text(face = "plain", color = "#000000", size = 10, angle = 0),
    legend.text = element_text(face = "plain", color = "#000000", size = 10, angle = 0)
  ) -> fitted.data
plot(fitted.data)

pdf(paste(expt, "_fitcurves_plot-noars-wide.pdf", sep = ""), width = 16, height = 4)
plot(fitted.data)
dev.off()
#fitted curves look satisfactory for continuing with analysis of AUC diffs
```

## plot without arsR, sd, only -Fe, 100uM Fe 
```{r}
stats.fin.2 <- filter (stats2, condition == c("(-Fe)", "(100Fe)"))

fin.cols<-c(pyrE="orangered4", arsR = "grey90", TroRHA = "medium blue", idr = "seagreen",  sirR ="gold", troR ="deepskyblue")

stats.fin.2 %>%
  ggplot(., aes(x = time, y = average, color = name, fill = name)) +
  xlab("time (h)") +
  ylab("log10(Absorbance at 600nm)") +
  geom_line(size = 1) +
  geom_ribbon(aes(ymin = average - sd, ymax = average + sd, fill = name, color = NULL), alpha = 0.2) +
  facet_wrap(~facet, nrow=1) +
  theme_classic() +
  #fill_color_manual(values = fin.cols) +
  scale_color_manual(values = fin.cols) +
  scale_fill_manual(values = fin.cols) +
  
  theme(
    plot.title = element_text(face = "bold", color = "#000000", size = 12, angle = 0),
    axis.title.x = element_text(face = "bold", color = "#000000", size = 10, angle = 0),
    axis.title.y = element_text(face = "bold", color = "#000000", size = 10, angle = 90),
    legend.title = element_blank(), legend.position = c(0.85, 0.15),
    axis.text.y = element_text(face = "plain", color = "#000000", size = 10, angle = 0),
    axis.text.x = element_text(face = "plain", color = "#000000", size = 10, angle = 0),
    legend.text = element_text(face = "plain", color = "#000000", size = 10, angle = 0)
  ) -> fitted.data
plot(fitted.data)

#pdf( "fitcurves_plot-fig3.pdf", width = 8, height = 4)
plot(fitted.data)
#dev.off()
#fitted curves look satisfactory for continuing with analysis of AUC diffs
```


#dtxR mutant growth parameters compared to WT 


## First calculate auc for WT and each dtx mutant
```{r}
colnames(d_gc)[1] <- "variable"
(merge(d_gc, mt0) -> tmp)
tmp$ID <- ifelse(tmp$condition=='',as.character(tmp$name),paste(tmp$name,' (',tmp$condition,')',sep=''))
tmp <- tmp[!(tmp$ID == ""), ]
tmp <- tmp[!(tmp$name == "blank"), ]

tmp %>%
  group_by(strain, condition, biorep, ID, name, media) %>%
 dplyr:: summarise(
    tgen = mean(t_gen),
    mu = mean(r),
    auc = mean(auc_l)
  ) %>% # average technical replicates
  group_by(condition, strain, name) %>% # average bioreps and calc sd
  dplyr::summarise(
    avg_tgen = mean(tgen),
    tgen_sd = sd(tgen),
    avg_mu = mean(mu),
    mu_sd = sd(mu),
    avg_auc = mean(auc),
    auc_sd = sd(auc)
  ) -> norms.dtx
```

```{r}
#write.output.files

write.csv (tmp, file = "exp4.growth.params.csv")
#tmp.fin <-tmp[!(tmp$condition =="Fe+ascorbate" | tmp$condition == "Fe+L-cys" | tmp$name == "ArsRHA"), ]
#write.csv(tmp.fin, file = "exp7.growth.params.fin.csv")
```

## Now plot auc for WT and each dtx mutant
```{r}
#  parameters compared to WT (by condition, not yet normalized to WT)

norms.dtx$facet = factor(norms.dtx$condition, levels = c("(--Fe)", "(-Fe)", "(1Fe)", "(10Fe)", "(100Fe)", "(250Fe)"))

norms.dtx %>%
  ggplot(aes(x = name, y = avg_auc, fill = name)) +
  xlab("") +
  ylab("AUC") +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9)) +
  geom_errorbar(aes(ymin = avg_auc - auc_sd, ymax = avg_auc + auc_sd),
    width = .1,
    size = 1, position = position_dodge(0.9)
  ) +
  facet_wrap(~facet, nrow = 1) +
  
  #scale_fill_viridis("strain", discrete = TRUE, begin = 0.2, end = 0.8, direction = -1) +
  scale_color_manual('strain description', values = rainbow(5)) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.02))) +
  theme_bw() +
  theme(
    axis.title.y = element_text(face = "plain", color = "#000000", size = 14, angle = 90),
    axis.text.y = element_text(face = "plain", color = "#000000", size = 12, angle = 0),
    axis.text.x = element_text(face = "plain", color = "#000000", size = 12, angle = 45, hjust = 1),
    panel.grid.major = element_blank(), panel.grid.minor = element_blank()
  ) -> figure

pdf(paste(expt, "_AUC_plot-facets.pdf", sep = ""), width = 16, height = 4)
plot(figure)
dev.off()

#can't include arsR in this figure - not enough data points to get error bars and stats. will need to have a separate sub-panel with arsR from a different experiment.
```

## AUC plot without arsR
```{r}
norms.dtx.noars <- norms.dtx[!(norms.dtx$name == 'arsR'), ]
norms.dtx.noars$facet = factor(norms.dtx.noars$condition, levels = c("(--Fe)", "(-Fe)", "(1Fe)", "(10Fe)", "(100Fe)", "(250Fe)"))
mycols<-  c(idr = "F8766D", pyrE="#7CAE00", sirR ="#00BFC4", troR ="#C77CFF")
  
norms.dtx.noars %>%
  ggplot(aes(x = name, y = avg_auc, fill = name)) +
  xlab("") +
  ylab("AUC") +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9)) +
  geom_errorbar(aes(ymin = avg_auc - auc_sd, ymax = avg_auc + auc_sd),
    width = .1,
    size = 1, position = position_dodge(0.9)
  ) +
  facet_wrap(~facet, nrow = 1) +
  
  #scale_fill_viridis("strain", discrete = TRUE, begin = 0.2, end = 0.8, direction = -1) +
  scale_color_manual(values = mycols) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.02))) +
  theme_bw() +
  theme(
    axis.title.y = element_text(face = "plain", color = "#000000", size = 14, angle = 90),
    axis.text.y = element_text(face = "plain", color = "#000000", size = 12, angle = 0),
    axis.text.x = element_text(face = "plain", color = "#000000", size = 12, angle = 45, hjust = 1),
    panel.grid.major = element_blank(), panel.grid.minor = element_blank()
  ) -> figure

pdf(paste(expt, "_AUC_plot-facets-noars.pdf", sep = ""), width = 16, height = 4)
plot(figure)
dev.off()

```


## plot td
```{r}
norms.dtx %>%
  ggplot(aes(x = name, y = avg_tgen, fill = name)) +
  xlab("") +
  ylab("generation time (hr)") +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9)) +
  geom_errorbar(aes(ymin = avg_tgen - tgen_sd, ymax = avg_tgen + tgen_sd),
    width = .1,
    size = 1, position = position_dodge(.9)
  ) +
  facet_wrap(~facet, nrow = 1) +
  #scale_fill_viridis("strain", discrete = TRUE, begin = 0.2, end = 0.8, direction = -1) +
  scale_color_manual('strain description', values = rainbow(5)) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.02))) +
  theme_bw() +
  theme(axis.title.y = element_text(face = "plain", color = "#000000", size = 14, angle = 90), axis.text.y = element_text(face = "plain", color = "#000000", size = 12, angle = 0), axis.text.x = element_text(face = "plain", color = "#000000", size = 12, angle = 45, hjust = 1), panel.grid.major = element_blank(), panel.grid.minor = element_blank()) -> figure

pdf(paste(expt, "_td_plot-facets.pdf", sep = ""), width = 16, height = 4)
plot(figure)
dev.off()
```

## plot tmu
```{r}
norms.dtx %>%
  ggplot(aes(x = name, y = avg_mu, fill = name)) +
  xlab("") +
  ylab(expression(mu)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9)) +
  geom_errorbar(aes(ymin = avg_mu - mu_sd, ymax = avg_mu + mu_sd),
    width = .1,
    size = 1, position = position_dodge(.9)
  ) +
  facet_wrap(~facet, nrow = 1) +
  #scale_fill_viridis("strain", discrete = TRUE, begin = 0.2, end = 0.8, direction = -1) +
  scale_color_manual('strain description', values = rainbow(5)) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.02))) +
  theme_bw() +
  theme(axis.title.y = element_text(face = "plain", color = "#000000", size = 14, angle = 90), axis.text.y = element_text(face = "plain", color = "#000000", size = 12, angle = 0), axis.text.x = element_text(face = "plain", color = "#000000", size = 12, angle = 45, hjust = 1), panel.grid.major = element_blank(), panel.grid.minor = element_blank()) -> figure

pdf(paste(expt, "_tmu_plot-facets.pdf", sep = ""), width = 16, height = 4)
plot(figure)
dev.off()
```

## plot tmu no arsr
```{r}
norms.dtx.noars %>%
  ggplot(aes(x = name, y = avg_mu, fill = name)) +
  xlab("") +
  ylab(expression(mu)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9)) +
  geom_errorbar(aes(ymin = avg_mu - mu_sd, ymax = avg_mu + mu_sd),
    width = .1,
    size = 1, position = position_dodge(.9)
  ) +
  facet_wrap(~facet, nrow = 1) +
  #scale_fill_viridis("strain", discrete = TRUE, begin = 0.2, end = 0.8, direction = -1) +
  scale_color_manual(values = mycols) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.02))) +
  theme_bw() +
  theme(axis.title.y = element_text(face = "plain", color = "#000000", size = 14, angle = 90), axis.text.y = element_text(face = "plain", color = "#000000", size = 12, angle = 0), axis.text.x = element_text(face = "plain", color = "#000000", size = 12, angle = 45, hjust = 1), panel.grid.major = element_blank(), panel.grid.minor = element_blank()) -> figure

pdf(paste(expt, "_tmu_plot-facets-noars.pdf", sep = ""), width = 16, height = 4)
plot(figure)
dev.off()
```
## test for significance of difference between WT and each mutant
```{r}
# Error bars overlap for all strains (WT, sirR, idr) for Td, mu except troR. Td plot difficult to interpret because troR "doubling time" swamps out rest. mu easier ot interpret. some differences between strains for auc. t-test troR vs pyrE each condition for mu, all dtx mutants vs pyrE for auc. easiest to do all-by-all t-test and then choose out these comparisons from the test result data frame.

pwt.mu <- pairwise_t_test(tmp, r ~ ID, p.adjust.method = "BH")
pwt.auc <- pairwise_t_test(tmp, auc_l ~ ID, p.adjust.method = "BH")

write.table (pwt.mu, file = "exp4_ttest_mu.txt", sep = "\t")
write.table (pwt.auc, file = "exp4_ttest_auc.txt", sep = "\t")
```


##################################################################################################### END



