---
title: "2023-02-16-troR-bioscreen-compiled"
author: "Amy Schmid"
date: "2/16/2023"
output: html_document
---
# For the calculation of final growth statistics across all bioscreen runs (exp3,4,5,6,7)
```{r}
# fresh start
rm(list = ls()) 
```

```{r}
# load required libraries
library(plyr)
library(tidyr) 
library(dplyr)
library(tibble)
library(viridis)
library(ggplot2)
library(tinytex)
library(latexpdf)
library(data.table)
library(growthrates)
library(gridExtra)
library(growthcurver)
library(factoextra)
library(Rfast)
library(rstatix)
library(ggpubr)
library(plotrix)
```

#load files
```{r}
#load each growth metrics output file. Bring together into two separate data frames: exp345 is from earlier experiments (mutants vs parent, titration of Fe). exp67 is from later experiments, show that troRHA is complementing troR mutant phenotype.
## from 2023-01-27 analysis of exp4
exp4<- read.csv("exp4.growth.params.csv")
## from 2023-02-14 analysis of exp3, 6, 7
exp4$exp <- "exp4"

exp3<- read.csv("exp3.params.noHM.csv")
exp3$exp <- "exp3"

exp6<- read.csv("exp6.growth.params.fin.csv")
exp6$exp <- "exp6"

exp7<- read.csv("exp7.growth.params.fin.csv")
exp7$exp <- "exp7"

## from 2023-02-15 analysis of exp5
exp5<-read.csv("exp5.growth.params.fin.nosid.csv")
exp5$exp <- "exp5"

## get all the condition names in the same format
exp345 <- bind_rows(exp3, exp4, exp5) %>%
  
  mutate(condition = replace(condition, condition == '-Fe', '1xnoFe')) %>%
  mutate (condition = replace(condition, condition == '--Fe', '2xnoFe')) %>% 
  mutate (condition = replace(condition, condition == '(-Fe)', '1xnoFe')) %>%
  mutate (condition = replace(condition, condition == '(--Fe)', '2xnoFe')) %>%
  mutate (condition = replace(condition, condition == '(1Fe)', '1Fe')) %>%
  mutate (condition = replace(condition, condition == '(10Fe)', '10Fe')) %>%
  mutate (condition = replace(condition, condition == '(100Fe)', '100Fe')) %>%
  mutate (condition = replace(condition, condition == '(250Fe)', '250Fe')) %>%
  mutate (condition = replace(condition, condition == '100microMFe', '100Fe')) %>%
  mutate (condition = replace(condition, condition == '1microMFe', '1Fe'))
  # make condition names consistent

## get all the condition names in the same format
exp67 <- bind_rows(exp6, exp7) %>%
  select(-X, -X.1, -X.2) %>% #remove columns with extraneous info not shared across experiments
  mutate(condition = replace(condition, condition == '-Fe', '1xnoFe')) %>%
  mutate (condition = replace(condition, condition == '--Fe', '2xnoFe')) %>% 
  mutate (condition = replace(condition, condition == '(-Fe)', '1xnoFe')) %>%
  mutate (condition = replace(condition, condition == '(--Fe)', '2xnoFe')) %>%
  mutate (condition = replace(condition, condition == '(1Fe)', '1Fe')) %>%
  mutate (condition = replace(condition, condition == '(10Fe)', '10Fe')) %>%
  mutate (condition = replace(condition, condition == '(100Fe)', '100Fe')) %>%
  mutate (condition = replace(condition, condition == '(250Fe)', '250Fe')) %>%
  mutate (condition = replace(condition, condition == '100microMFe', '100Fe')) %>%
  mutate (condition = replace(condition, condition == '1microMFe', '1Fe'))
  # make condition names consistent

```


```{r}
#load each fitcurves output file for mutants and bring them together into two dataframes: c.exp345, c.exp67
## from 2023-01-27 analysis of exp4
exp4.c <- read.csv("exp4_fitcurves.csv")
## from 2023-02-14 analysis of exp3, 6, 7
exp4.c$exp <- "exp4"

exp3.c <- read.csv("~/Dropbox (Duke Bio_Ea)/projects/metal_project/HVO_dtxR/final-data/Bioscreen//data/exp3_fitcurves-parent.csv")
exp3.c$exp <- "exp3"

exp6.c <- read.csv("exp6_fitcurves.csv")
exp6.c$exp <- "exp6"

exp7.c <- read.csv("exp7_fitcurves.csv")
## from 2023-02-15 analysis of exp5
exp7.c$exp <- "exp7"

exp5.c <-read.csv("exp5_fitcurves.csv")
exp5.c$exp <- "exp5"

## get all the condition names in the same format
c.exp345 <- bind_rows(exp3.c, exp4.c, exp5.c) %>%
  #select(-X, -X.1, -X.2, -X.3, -X.4) %>% #remove columns with extraneous info not shared across experiments
  mutate(condition = replace(condition, condition == '-Fe', '1xnoFe')) %>%
  mutate (condition = replace(condition, condition == '--Fe', '2xnoFe')) %>% 
  mutate (condition = replace(condition, condition == '(-Fe)', '1xnoFe')) %>%
  mutate (condition = replace(condition, condition == '(--Fe)', '2xnoFe')) %>%
  mutate (condition = replace(condition, condition == '(1Fe)', '1Fe')) %>%
  mutate (condition = replace(condition, condition == '(10Fe)', '10Fe')) %>%
  mutate (condition = replace(condition, condition == '(100Fe)', '100Fe')) %>%
  mutate (condition = replace(condition, condition == '(250Fe)', '250Fe')) %>%
  mutate (condition = replace(condition, condition == '100microMFe', '100Fe')) %>%
  mutate (condition = replace(condition, condition == '1microMFe', '1Fe'))
  # make condition names consistent

## get all the condition names in the same format
c.exp67 <- bind_rows(exp6.c, exp7.c) %>%
  #select(-X, -X.1, -X.2, -X.3, -X.4) %>% #remove columns with extraneous info not shared across experiments
  mutate(condition = replace(condition, condition == '-Fe', '1xnoFe')) %>%
  mutate (condition = replace(condition, condition == '--Fe', '2xnoFe')) %>% 
  mutate (condition = replace(condition, condition == '(-Fe)', '1xnoFe')) %>%
  mutate (condition = replace(condition, condition == '(--Fe)', '2xnoFe')) %>%
  mutate (condition = replace(condition, condition == '(1Fe)', '1Fe')) %>%
  mutate (condition = replace(condition, condition == '(10Fe)', '10Fe')) %>%
  mutate (condition = replace(condition, condition == '(100Fe)', '100Fe')) %>%
  mutate (condition = replace(condition, condition == '(250Fe)', '250Fe')) %>%
  mutate (condition = replace(condition, condition == '100microMFe', '100Fe')) %>%
  mutate (condition = replace(condition, condition == '1microMFe', '1Fe'))
  # make condition names consistent

#filter out all conditions except 100Fe and -Fe for final plot of mutants, given that pyrE titration showed 100Fe as condition of choice
cfin.exp345 <- c.exp345[!(c.exp345$condition == "2xnoFe" | c.exp345$condition == "1Fe" | c.exp345$condition == "10Fe" | c.exp345$condition == "250Fe"), ]

cfin.exp67 <- c.exp67[!(c.exp67$condition == "2xnoFe" | c.exp67$condition == "1Fe" | c.exp67$condition == "10Fe" | c.exp67$condition == "250Fe"), ]

```


```{r}
#load each fitcurves output file for pyrE by itself and bring them together (exp3,4,5) for the purpose of plotting titration average curves
## from 2023-01-27 analysis of exp4
exp4.cp <- read.csv("exp4_fitcurves_parent.csv") 

exp4.cp$exp <- "exp4"

## from 2023-02-14 analysis of exp3, 6, 7
exp3.cp <- read.csv("exp3_fitcurves-parent.csv")

exp3.cp$exp <- "exp3"
#exp6.cp <- read.csv("exp6_fitcurves_parent.csv")

#exp6.cp$exp <- "exp6"

#exp7.cp <- read.csv("~/Dropbox (Duke #Bio_Ea)/projects/metal_project/HVO_dtxR/final-data/Bioscreen/data/exp7_fitcurves_parent.csv")

#exp7.cp$exp <- "exp7"

## from 2023-02-15 analysis of exp5
exp5.cp <-read.csv("exp5_fitcurves_parent.csv")
exp5.cp$exp <- "exp5"

## get all the condition names in the same format
all.cp <- bind_rows(exp3.cp, exp4.cp, exp5.cp) %>%
  #select(-X, -X.1, -X.2, -X.3, -X.4) %>% #remove columns with extraneous info not shared across experiments
  mutate(condition = replace(condition, condition == '-Fe', '1xnoFe')) %>%
  mutate (condition = replace(condition, condition == '--Fe', '2xnoFe')) %>% 
  mutate (condition = replace(condition, condition == '(-Fe)', '1xnoFe')) %>%
  mutate (condition = replace(condition, condition == '(--Fe)', '2xnoFe')) %>%
  mutate (condition = replace(condition, condition == '(1Fe)', '1Fe')) %>%
  mutate (condition = replace(condition, condition == '(10Fe)', '10Fe')) %>%
  mutate (condition = replace(condition, condition == '(100Fe)', '100Fe')) %>%
  mutate (condition = replace(condition, condition == '(250Fe)', '250Fe')) %>%
  mutate (condition = replace(condition, condition == '100microMFe', '100Fe')) %>%
  mutate (condition = replace(condition, condition == '1microMFe', '1Fe'))
  # make condition names consistent


```

# plot line graphs, avg curves pyrE across conditions
```{r}
#first get summary stats
titr.cp <- all.cp %>% #filter(all.cp, name == "pyrE" ) %>% 
  #filter (! condition == "1Fe") %>%
 # group_by(strain, condition, name) %>%
 #dplyr:: summarise(
#    tgen = mean(t_gen),
#    mu = mean(r),
 #   auc = mean(auc_l)
#  ) %>% # average technical replicates
  group_by(condition, strain, name, time) %>% # average bioreps and calc sd
  dplyr::summarise(
    avg_fits = mean(log10(average + 1)),
    sd_fits = sd(log10(sd + 1)),
    sem_fits = std.error(log10(sem + 1))
  ) -> fits.pyr #make a dataframe, choosing the rows that contain the parent strain.

```

#plot line graph, pyrE titration
```{r}
#fits.pyr <- filter (stats2, name == "pyrE")
cond.cols <- c(`2xnoFe` = "bisque",  `1xnoFe` = "bisque3", `1Fe` = "pink", `10Fe` = "orangered3", `100Fe` = "orangered4", `250Fe` = "black")
#stats2.titr$facet = factor(stats2.titr$condition, levels = c("(--Fe)", "(-Fe)", "(10Fe)", "(100Fe)", "(250Fe)"))

#remove time points not consistent across datasets with no sd (not enough data)
#fits.pyr <- fits.pyr %>% 
#filter(!row_number() %in% c(696,697))

fits.pyr %>%
  ggplot(., aes(x = time, y = avg_fits, color = condition)) +
  xlab("time (h)") +
  ylab("log10(Absorbance at 600nm)") +
  geom_line(size = 1) +
 # geom_ribbon(aes(ymin = avg_fits - sem_fits, ymax = avg_fits + sem_fits, fill = condition), alpha = 0.2) +
 # facet_wrap(~exp, nrow = 1) +
  
  theme_classic() +
  # scale_color_manual('strain description', values = rainbow(5)) +
  scale_color_manual ('condition', values = cond.cols) +
  scale_fill_manual("condition", values = cond.cols) +
  theme(
    plot.title = element_text(face = "bold", color = "#000000", size = 12, angle = 0),
    axis.title.x = element_text(face = "bold", color = "#000000", size = 10, angle = 0),
    axis.title.y = element_text(face = "bold", color = "#000000", size = 10, angle = 90),
    legend.title = element_blank(), legend.position = c(0.85, 0.15),
    axis.text.y = element_text(face = "plain", color = "#000000", size = 10, angle = 0),
    axis.text.x = element_text(face = "plain", color = "#000000", size = 10, angle = 0),
    legend.text = element_text(face = "plain", color = "#000000", size = 10, angle = 0)
  ) -> fitted.data.avgs
plot(fitted.data.avgs)

#pdf("trial-compiled_fitcurves_pyre-titr-plots.pdf", height = 4, width = 6) 
plot(fitted.data.avgs)
#dev.off()

#1Fe suppports faster growth than the other conditions, but has lower A. only 6 trials, included in only 1 experiment. Leave out 1Fe since it is not same n as other conditions.
```

# Final plot (Fig 1A), no 1Fe
```{r}
fits.pyr.no1<- fits.pyr[!(fits.pyr$condition == "1Fe"), ]
#fits.pyr <- filter (stats2, name == "pyrE")
cond.cols <- c(`2xnoFe` = "bisque",  `1xnoFe` = "pink", `10Fe` = "orangered3", `100Fe` = "orangered4", `250Fe` = "black")
#stats2.titr$facet = factor(stats2.titr$condition, levels = c("(--Fe)", "(-Fe)", "(10Fe)", "(100Fe)", "(250Fe)"))

#remove time points not consistent across datasets with no sd (not enough data)
#fits.pyr <- fits.pyr %>% 
#filter(!row_number() %in% c(696,697))

fits.pyr.no1 %>%
  ggplot(., aes(x = time, y = avg_fits, color = condition)) +
  xlab("time (h)") +
  ylab("log10(Absorbance at 600nm)") +
  geom_line(size = 1) +
 # geom_ribbon(aes(ymin = avg_fits - sem_fits, ymax = avg_fits + sem_fits, fill = condition), alpha = 0.2) +
 # facet_wrap(~exp, nrow = 1) +
  
  theme_classic() +
  # scale_color_manual('strain description', values = rainbow(5)) +
  scale_color_manual ('condition', values = cond.cols) +
  scale_fill_manual("condition", values = cond.cols) +
  theme(
    plot.title = element_text(face = "bold", color = "#000000", size = 12, angle = 0),
    axis.title.x = element_text(face = "bold", color = "#000000", size = 10, angle = 0),
    axis.title.y = element_text(face = "bold", color = "#000000", size = 10, angle = 90),
    legend.title = element_blank(), legend.position = c(0.85, 0.15),
    axis.text.y = element_text(face = "plain", color = "#000000", size = 10, angle = 0),
    axis.text.x = element_text(face = "plain", color = "#000000", size = 10, angle = 0),
    legend.text = element_text(face = "plain", color = "#000000", size = 10, angle = 0)
  ) -> fitted.data.avgs.no1
plot(fitted.data.avgs.no1)

#pdf("Fig1A-compiled_fitcurves_pyre-titr-plots.pdf", height = 4, width = 6) ## FINAL fig 1a. uncomment to print final plot to file.
plot(fitted.data.avgs.no1)
#dev.off()


```

# plot line graphs, avg curves mutants
```{r}
#first get summary stats
titr.c <- c.exp345 %>% #filter(all.cp, name == "pyrE" ) %>% 
  #filter (! condition == "1Fe") %>%
 # group_by(strain, condition, name) %>%
 #dplyr:: summarise(
#    tgen = mean(t_gen),
#    mu = mean(r),
 #   auc = mean(auc_l)
#  ) %>% # average technical replicates
  group_by(condition, strain, name, time) %>% # average bioreps and calc sd
  dplyr::summarise(
    avg_fits = mean(average),
    sd_fits = sd(average),
    sem_fits = std.error(average)
  ) -> fits.dtx #make a dataframe, choosing the rows that contain the parent strain.

```

#plot line graph, dtx
```{r}
my.cols<-  c(pyrE="orangered4", arsR = "grey90",  idr = "seagreen",  sirR ="gold", troR ="deepskyblue")
#TroRHA = "medium blue",

#focus on 1Fe and 1xnoFe
fits.dtx.fin <- filter (fits.dtx, condition == c("1xnoFe", "100Fe") )
fits.dtx.fin$facet = factor(fits.dtx.fin$condition, levels = c("1xnoFe", "100Fe"))
fits.dtx.fin %>%
  ggplot(., aes(x = time, y = avg_fits, color = name)) +
  xlab("time (h)") +
  ylab("log10(Absorbance at 600nm)") +
  geom_line(size = 1) +
 # geom_ribbon(aes(ymin = average - sem, ymax = average + sem, fill = name, color = NULL), alpha = 0.2) +
  facet_wrap(~facet, nrow = 1) +
  
  theme_classic() +
  # scale_color_manual('strain description', values = rainbow(5)) +
  scale_color_manual ('strain', values = my.cols) +
 # scale_fill_manual("condition", values = cond.cols) +
  theme(
    plot.title = element_text(face = "bold", color = "#000000", size = 12, angle = 0),
    axis.title.x = element_text(face = "bold", color = "#000000", size = 10, angle = 0),
    axis.title.y = element_text(face = "bold", color = "#000000", size = 10, angle = 90),
    legend.title = element_blank(), legend.position = c(0.85, 0.15),
    axis.text.y = element_text(face = "plain", color = "#000000", size = 10, angle = 0),
    axis.text.x = element_text(face = "plain", color = "#000000", size = 10, angle = 0),
    legend.text = element_text(face = "plain", color = "#000000", size = 10, angle = 0)
  ) -> fitted.data.dtx.avgs
plot(fitted.data.dtx.avgs)

#pdf("Fig3A-compiled_fitcurves_dtx-plots.pdf", height = 4, width = 6)
plot(fitted.data.dtx.avgs)
#dev.off()


```

```{r}
#first get summary stats by experiment
c.exp345 %>% #filter(all.cp, name == "pyrE" ) %>% 
  #filter (! condition == "1Fe") %>%
 # group_by(strain, condition, name) %>%
 #dplyr:: summarise(
#    tgen = mean(t_gen),
#    mu = mean(r),
 #   auc = mean(auc_l)
#  ) %>% # average technical replicates
  group_by(condition, strain, name, time, exp) %>% # average bioreps and calc sd
  dplyr::summarise(
    avg_fits = mean(average),
    sd_fits = sd(average),
    sem_fits = std.error(average)
  ) -> fits.dtx.exp #make a dataframe, choosing the rows that contain the parent strain.
```

```{r}
#plot facet grid, by experiment and condition to compare all together
fits.dtx.exp.fin <- fits.dtx.exp  %>%
  filter (condition == "1xnoFe" | condition == "100Fe")

fits.dtx.exp.fin %>%
  ggplot(., aes(x = time, y = avg_fits, color = name)) +
  xlab("time (h)") +
  ylab("log10(Absorbance at 600nm)") +
  geom_line(size = 1) +
 # geom_ribbon(aes(ymin = average - sem, ymax = average + sem, fill = name, color = NULL), alpha = 0.2) +
  facet_grid(condition~exp) +
  
  theme_classic() +
  # scale_color_manual('strain description', values = rainbow(5)) +
  scale_color_manual ('strain', values = my.cols) +
 # scale_fill_manual("condition", values = cond.cols) +
  theme(
    plot.title = element_text(face = "bold", color = "#000000", size = 12, angle = 0),
    axis.title.x = element_text(face = "bold", color = "#000000", size = 10, angle = 0),
    axis.title.y = element_text(face = "bold", color = "#000000", size = 10, angle = 90),
    legend.title = element_blank(), legend.position = c(0.85, 0.15),
    axis.text.y = element_text(face = "plain", color = "#000000", size = 10, angle = 0),
    axis.text.x = element_text(face = "plain", color = "#000000", size = 10, angle = 0),
    legend.text = element_text(face = "plain", color = "#000000", size = 10, angle = 0)
  ) -> fitted.data.dtx.all
plot(fitted.data.dtx.all)

#pdf("all_compiled_fitcurves_dtx-plots_bycond_exp.pdf", height = 4, width = 6)
plot(fitted.data.dtx.all)
#dev.off()

```




# plot bar graph of AUC, pyrE only titration 
```{r}
#first focus the data frame on pyrE only and get summary stats
titr <- filter(exp345, name == "pyrE" ) %>% 
 # filter (! condition == "1Fe") %>%
 # filter (! condition == "2xnoFe") %>% #2xnoFe doesn't grow, so doesn't make sense to plot "growth metrics"
 # group_by(strain, condition, name) %>%
 #dplyr:: summarise(
#    tgen = mean(t_gen),
#    mu = mean(r),
 #   auc = mean(auc_l)
#  ) %>% # average technical replicates
  group_by(condition, strain, name) %>% # average bioreps and calc sd
  dplyr::summarise(
    avg_tgen = mean(t_gen),
    tgen_sd = sd(t_gen),
    avg_mu = mean(r),
    mu_sd = sd(r),
    mu_sem = std.error(r),
    avg_auc = mean(auc_l),
    auc_sd = sd(auc_l),
    auc_sem = std.error(auc_l)
  ) -> params.pyr #make a dataframe, choosing the rows that contain the parent strain.
```
## preliminary plot AUC bar chart with sd as error bars 
```{r}
params.pyr$reord = factor(params.pyr$condition, levels = c("2xnoFe", "1xnoFe" ,  "1Fe", "10Fe", "100Fe", "250Fe")) #put the conditions in the right order for plotting


#"2xnoFE" does not grow, AUC and mu will be thrown off, take out

#cond.cols <- c(`1xnoFe` = "bisque", `1Fe` = "orangered",  `10Fe` = "orangered3", `100Fe` = "orangered4", `250Fe` = "black")


params.pyr %>%
  ggplot(aes(x = reord, y = avg_auc, fill = condition)) +
  xlab("") +
  ylab("AUC") +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9)) +
  geom_errorbar(aes(ymin = avg_auc - auc_sem, ymax = avg_auc + auc_sem),
    width = .1,
    size = 1, position = position_dodge(0.9)
  ) +
#  facet_wrap(~facet, nrow = 1) +
  
  #scale_fill_viridis("strain", discrete = TRUE, begin = 0.2, end = 0.8, direction = -1) +
  scale_fill_manual("[Fe], micromolar", values = cond.cols) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.02))) +
  theme_bw() +
  theme(
    axis.title.y = element_text(face = "plain", color = "#000000", size = 14, angle = 90),
    axis.text.y = element_text(face = "plain", color = "#000000", size = 12, angle = 0),
    axis.text.x = element_text(face = "plain", color = "#000000", size = 12, angle = 45, hjust = 1),
    panel.grid.major = element_blank(), panel.grid.minor = element_blank()
  ) -> figure

#pdf( "trial-AUC_plot-titration-exp345.pdf", width = 6, height = 4)
plot(figure)
#dev.off()

#why is 1xnoFe AUC higher than that of 1Fe?  plot also mu - according to curves it looks like 1Fe grew fast but then quickly hit stationary phase? Could also be due to low n for 1Fe (6) relative to other conditions (9-12).
```

## Final plot bar chart with sd as error bars (Fig 1B)
```{r}
params.pyr$reord = factor(params.pyr$condition, levels = c("2xnoFe", "1xnoFe" ,  "10Fe", "100Fe", "250Fe")) #put the conditions in the right order for plotting


#"2xnoFE" does not grow, AUC and mu will be thrown off, take out

#cond.cols <- c(`1xnoFe` = "bisque", `1Fe` = "orangered",  `10Fe` = "orangered3", `100Fe` = "orangered4", `250Fe` = "black")
params.pyr.no1<- params.pyr[!(params.pyr$condition == "1Fe"), ]

params.pyr.no1 %>%
  ggplot(aes(x = reord, y = avg_auc, fill = condition)) +
  xlab("") +
  ylab("AUC") +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9)) +
  geom_errorbar(aes(ymin = avg_auc - auc_sem, ymax = avg_auc + auc_sem),
    width = .1,
    size = 1, position = position_dodge(0.9)
  ) +
#  facet_wrap(~facet, nrow = 1) +
  
  #scale_fill_viridis("strain", discrete = TRUE, begin = 0.2, end = 0.8, direction = -1) +
  scale_fill_manual("[Fe], micromolar", values = cond.cols) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.02))) +
  theme_bw() +
  theme(
    axis.title.y = element_text(face = "plain", color = "#000000", size = 14, angle = 90),
    axis.text.y = element_text(face = "plain", color = "#000000", size = 12, angle = 0),
    axis.text.x = element_text(face = "plain", color = "#000000", size = 12, angle = 45, hjust = 1),
    panel.grid.major = element_blank(), panel.grid.minor = element_blank()
  ) -> figure

#pdf( "Fig1B-AUC_plot-titration-exp345.pdf", width = 6, height = 4)
plot(figure)
#dev.off()


```





#preliminary plot of mu for pyrE titration
```{r}
params.pyr %>%
  ggplot(aes(x = reord, y = avg_mu, fill = condition)) +
  xlab("") +
  ylab("mu") +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9)) +
  geom_errorbar(aes(ymin = avg_mu - mu_sem, ymax = avg_mu + mu_sem),
    width = .1,
    size = 1, position = position_dodge(0.9)
  ) +
#  facet_wrap(~facet, nrow = 1) +
  
  #scale_fill_viridis("strain", discrete = TRUE, begin = 0.2, end = 0.8, direction = -1) +
  scale_fill_manual("[Fe], micromolar", values = cond.cols) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.02))) +
  theme_bw() +
  theme(
    axis.title.y = element_text(face = "plain", color = "#000000", size = 14, angle = 90),
    axis.text.y = element_text(face = "plain", color = "#000000", size = 12, angle = 0),
    axis.text.x = element_text(face = "plain", color = "#000000", size = 12, angle = 45, hjust = 1),
    panel.grid.major = element_blank(), panel.grid.minor = element_blank()
  ) -> figure

#pdf( "trial-mu_plot-titration-exp345.pdf", width = 6, height = 4)
plot(figure)
#dev.off()

#yes - 1Fe has much faster mu than no fe conditions and 10Fe. Suggest removal of 1Fe.
```
# final figure, mu
```{r}
params.pyr.no1 %>%
  ggplot(aes(x = reord, y = avg_mu, fill = condition)) +
  xlab("") +
  ylab("mu") +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9)) +
  geom_errorbar(aes(ymin = avg_mu - mu_sem, ymax = avg_mu + mu_sem),
    width = .1,
    size = 1, position = position_dodge(0.9)
  ) +
#  facet_wrap(~facet, nrow = 1) +
  
  #scale_fill_viridis("strain", discrete = TRUE, begin = 0.2, end = 0.8, direction = -1) +
  scale_fill_manual("[Fe], micromolar", values = cond.cols) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.02))) +
  theme_bw() +
  theme(
    axis.title.y = element_text(face = "plain", color = "#000000", size = 14, angle = 90),
    axis.text.y = element_text(face = "plain", color = "#000000", size = 12, angle = 0),
    axis.text.x = element_text(face = "plain", color = "#000000", size = 12, angle = 45, hjust = 1),
    panel.grid.major = element_blank(), panel.grid.minor = element_blank()
  ) -> figure

#pdf( "Fig1C-mu_plot-titration-exp345.pdf", width = 6, height = 4)
plot(figure)
#dev.off()

#==
```
# Stats, Fe titration in pyrE with 1Fe
```{r}
titr.fin <- filter(exp345, name == "pyrE" )# %>% 
  #filter (! condition == "1Fe") 

pwt.mu <- pairwise_t_test(titr.fin, r ~ condition, p.adjust.method = "BH")
pwt.auc <- pairwise_t_test(titr.fin, auc_l ~ condition, p.adjust.method = "BH")

write.csv (pwt.mu, file = "exp345_titr_ttest_mu.csv")
write.csv (pwt.auc, file = "exp345_titr_ttest_auc.csv")

```

# Stats, Fe titration in pyrE without 1Fe
```{r}
titr.fin <- filter(exp345, name == "pyrE" ) %>% 
  filter (! condition == "1Fe") 

pwt.mu <- pairwise_t_test(titr.fin, r ~ condition, p.adjust.method = "BH")
pwt.auc <- pairwise_t_test(titr.fin, auc_l ~ condition, p.adjust.method = "BH")

write.csv (pwt.mu, file = "exp345_titr_ttest_mu-no1Fe.csv")
write.csv (pwt.auc, file = "exp345_titr_ttest_auc-no1Fe.csv")

```

# Plot bar graph of AUC,mutants plus pyrE (Fig 3c)
## Get the data frame ready
```{r}
#first focus the data frame on 1xnoFe and 100Fe only and get summary stats
m <- exp345[!(exp345$condition == "2xnoFe" | exp345$condition == "1Fe" | exp345$condition == "10Fe" | exp345$condition == "250Fe"), ] %>%
 # group_by(strain, condition, name) %>%
 #dplyr:: summarise(
#    tgen = mean(t_gen),
#    mu = mean(r),
 #   auc = mean(auc_l)
#  ) %>% # average technical replicates
  group_by(condition, strain, name) %>% # average bioreps and calc sd
  dplyr::summarise(
    avg_tgen = mean(t_gen),
    tgen_sd = sd(t_gen),
    avg_mu = mean(r),
    mu_sd = sd(r),
    mu_sem = std.error(r),
    avg_auc = mean(auc_l),
    auc_sd = sd(auc_l),
    auc_sem = std.error(auc_l)
  ) -> params.m #make a dataframe, choosing the rows that contain the parent strain.


```
## plot the AUC bar graph
```{r}
# set the color scheme
my.cols<-  c(pyrE="orangered4", arsR = "grey90", idr = "seagreen",  sirR ="gold", troR ="deepskyblue")
#TroRHA = "medium blue", 

#order the bars for faceting
params.m$facet = factor(params.m$condition, levels = c("1xnoFe", "100Fe"))
params.m$ordering = factor (params.m$name, levels = c("pyrE", "arsR", "idr", "sirR", "troR"))

params.m %>%
  
  ggplot(aes(x = ordering, y = avg_auc, fill = name)) +
  xlab("") +
  ylab("AUC") +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9)) +
  geom_errorbar(aes(ymin = avg_auc - auc_sem, ymax = avg_auc + auc_sem),
    width = .1,
    size = 1, position = position_dodge(0.9)
  ) +
  facet_wrap(~facet, nrow = 1) +
  
  #scale_fill_viridis("strain", discrete = TRUE, begin = 0.2, end = 0.8, direction = -1) +
  scale_fill_manual("strain", values = my.cols) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.02))) +
  theme_bw() +
  theme(
    axis.title.y = element_text(face = "plain", color = "#000000", size = 14, angle = 90),
    axis.text.y = element_text(face = "plain", color = "#000000", size = 12, angle = 0),
    axis.text.x = element_text(face = "plain", color = "#000000", size = 12, angle = 45, hjust = 1),
    panel.grid.major = element_blank(), panel.grid.minor = element_blank()
  ) -> figure

pdf( "Fig3B-exp345-AUC_plot-dtx.pdf", width = 6, height = 4)
plot(figure)
dev.off()

```

## plot the mu bar graph, fig3d
```{r}
# set the color scheme
my.cols<-  c(pyrE="orangered4", arsR = "grey90",idr = "seagreen",  sirR ="gold", troR ="deepskyblue")
# TroRHA = "medium blue", 

#order the bars for faceting
params.m$facet = factor(params.m$condition, levels = c("1xnoFe", "100Fe"))
params.m$ordering = factor (params.m$name, levels = c("pyrE", "arsR", "idr", "sirR", "troR"))

params.m %>%
  
  ggplot(aes(x = ordering, y = avg_mu, fill = name)) +
  xlab("") +
  ylab("mu") +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9)) +
  geom_errorbar(aes(ymin = avg_mu - mu_sem, ymax = avg_mu + mu_sem),
    width = .1,
    size = 1, position = position_dodge(0.9)
  ) +
  facet_wrap(~facet, nrow = 1) +
  
  #scale_fill_viridis("strain", discrete = TRUE, begin = 0.2, end = 0.8, direction = -1) +
  scale_fill_manual("strain", values = my.cols) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.02))) +
  theme_bw() +
  theme(
    axis.title.y = element_text(face = "plain", color = "#000000", size = 14, angle = 90),
    axis.text.y = element_text(face = "plain", color = "#000000", size = 12, angle = 0),
    axis.text.x = element_text(face = "plain", color = "#000000", size = 12, angle = 45, hjust = 1),
    panel.grid.major = element_blank(), panel.grid.minor = element_blank()
  ) -> figure

pdf( "Fig3C-mu_plot-dtx-exp345.pdf", width = 6, height = 4)
plot(figure)
dev.off()

```

# Stats, dtx mutants all by all
```{r}
dtx.fin <- filter(exp345, condition == "1xnoFe" | condition == "100Fe" ) 
  #filter (! condition == "1Fe") 

dtx.fin.noFe<- filter (dtx.fin, condition == "1xnoFe")
dtx.fin.Fe <- filter (dtx.fin, condition == "100Fe")

pwt.mu.noFe <- pairwise_t_test(dtx.fin.noFe, r ~ name, p.adjust.method = "BH")
pwt.mu.Fe <- pairwise_t_test(dtx.fin.Fe, r~name, p.adjust.method = "BH")
pwt.auc.noFe <- pairwise_t_test(dtx.fin.noFe, auc_l ~ name, p.adjust.method = "BH")
pwt.auc.Fe <- pairwise_t_test (dtx.fin.Fe, auc_l ~name, p.adjust.method = "BH")

write.csv (pwt.mu.noFe, file = "exp345_dtx.ttest_mu-noFe.csv")
write.csv (pwt.mu.Fe, file = "exp345_dtx_Fe.ttest_mu-Fe.csv")
write.csv (pwt.auc.noFe, file = "exp345_dtx_ttest_auc-noFe.csv")
write.csv (pwt.auc.Fe, file = "exp345_dtx_ttext_auc-Fe.csv")

```
